High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-05-09 notebooks/cal_housing.qmd
added lin sep
more results

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks)
```

@@

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(λ=λ₁),
    "REVISE" => REVISEGenerator(λ=λ₁),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
)
```

@@

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="California Housing",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
    converge_when=:generator_conditions,
)
CSV.write(joinpath(output_path, "cal_housing_benchmark.csv"), bmk())
```

```{julia}
@chain bmk() begin
    @group_by(dataname, generator, model, variable)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable == "distance_from_energy")
end
```

>>>>

```{julia}
include("$(pwd())/notebooks/setup.jl")
eval(setup_notebooks)
```

@@

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(),
    "REVISE" => REVISEGenerator(),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
)
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="California Housing",
            n_individuals=10,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "cal_housing_benchmark.csv"), bmk())
```

### 2023-05-09 notebooks/circles.qmd
more results

```{julia}
λ₁ = 0.25
λ₂ = 0.75
λ₃ = 0.75
Λ = [λ₁, λ₂, λ₃]

@@

```{julia}
M = model_dict["JEM"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[1]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[2]
factual = predict_label(M, counterfactual_data, x_factual)[1]

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(ce, title=name, alpha=0.2, cbar=false, axis=nothing)
    if name == "ECCCo"
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.5
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(500,520))
display(plt)
savefig(plt, joinpath(output_images_path, "circles_poc.png"))
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="Circles",
            n_individuals=5,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "circles_benchmark.csv"), bmk())
```

>>>>

```{julia}
λ₁ = 0.1
λ₂ = 0.25
λ₃ = 0.25
Λ = [λ₁, λ₂, λ₃]

@@

```{julia}
Random.seed!(2023)

M = model_dict["JEM"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[2]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[1]
factual = predict_label(M, counterfactual_data, x_factual)[1]

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(ce, title=name, alpha=0.2, cbar=false, axis=nothing)
    if contains(name, "ECCCo")
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.5
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(650,500))
display(plt)
savefig(plt, joinpath(output_images_path, "circles_poc.png"))
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="Circles",
            n_individuals=25,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "circles_benchmark.csv"), bmk())
```

### 2023-05-09 notebooks/gmsc.qmd
more results

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,5e-2]      # penalty strengths
```

@@

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(λ=λ₁),
    "REVISE" => REVISEGenerator(λ=λ₁),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
)
```

```{julia}
# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.Evaluation.validity,
    CounterfactualExplanations.Evaluation.redundancy,
]

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="GMSC",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
)
CSV.write(joinpath(output_path, "gmsc_benchmark.csv"), bmk())
```

```{julia}
@chain bmk() begin
    @group_by(dataname, generator, model, variable)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable == "distance_from_energy")
end
```

>>>>

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=50,
)
α = [1.0,1.0,1e-2]      # penalty strengths
```

@@

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(),
    "REVISE" => REVISEGenerator(),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
)
```

```{julia}
# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.Evaluation.validity,
    CounterfactualExplanations.Evaluation.redundancy,
    ECCCo.set_size_penalty
]

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="GMSC",
            n_individuals=10,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "gmsc_benchmark.csv"), bmk())
```

### 2023-05-09 notebooks/linearly_separable.qmd
more results

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks)
```

# Circles Data

@@

# Data:
test_size = 0.2
n_obs = Int(1000 / (1.0 - test_size))
counterfactual_data, test_data = train_test_split(load_linearly_separable(n_obs); test_size=test_size)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
output_dim = length(unique(labels))
```

@@

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=batch_size,
)
α = [1.0,1.0,1e-2]      # penalty strengths
```

@@

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(model; mod_name=mod_name) for (mod_name, model) in models)
    Serialization.serialize(joinpath(output_path,"circles_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"circles_models.jls"))
end
```

@@

measure = Dict(
    :f1score => multiclass_f1score, 
    :acc => accuracy, 
    :precision => multiclass_precision
)
model_performance = DataFrame()
for (mod_name, model) in model_dict
    # Test performance:
    _perf = CounterfactualExplanations.Models.model_evaluation(model, test_data, measure=collect(values(measure)))
    _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
    _perf.mod_name .= mod_name
    model_performance = vcat(model_performance, _perf)
end
Serialization.serialize(joinpath(output_path,"circles_model_performance.jls"), model_performance)
CSV.write(joinpath(output_path, "circles_model_performance.csv"), model_performance)
model_performance
```

```{julia}
n_regen = 200
n_each = batch_size
for (mod_name, model) in model_dict
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
    opt = ImproperSGLD()
    plts = []
    for target in levels(labels)
        target_idx = findall(levels(labels) .== target)[1]
        f(x) = logits(model, x)
        X̂ = sampler(f, opt; niter=n_regen, n_samples=n_each, y=target_idx)
        ex = extrema(hcat(MLJFlux.reformat(X),X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=100)
        x2 = range(1.0f0.*ylims...,length=100)
        p(x) = probs(model, x)
        plt = Plots.contour(
            x1, x2, (x, y) -> p([x, y][:,:])[target_idx], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(
            MLJFlux.reformat(X)[1,:], MLJFlux.reformat(X)[2,:], 
            color=Int.(labels.refs).-1, group=Int.(labels.refs).-1, alpha=0.5
        )
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target], size(X̂,2)), 
            group=repeat([target], size(X̂,2)), 
            shape=:star5, ms=10
        )
        savefig(plt, joinpath(output_images_path, "circles_generated_$(mod_name).png"))
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400), plot_title=mod_name)
    display(plt)
end
```

@@

```{julia}
M = model_dict["JEM"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[1]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[2]
factual = predict_label(M, counterfactual_data, x_factual)[1]

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(ce, title=name, alpha=0.2, cbar=false, axis=nothing)
    if name == "ECCCo"
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.5
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(500,520))
display(plt)
savefig(plt, joinpath(output_images_path, "circles_poc.png"))
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="Circles",
            n_individuals=5,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "circles_benchmark.csv"), bmk())
```

```{julia}
df = @chain bmk() begin
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
    @mutate(variable = ifelse.(variable .== "distance", "Cost", variable))
    @mutate(variable = ifelse.(variable .== "redundancy", "Redundancy", variable))
    @mutate(variable = ifelse.(variable .== "Validity", "Validity", variable))
end
plt = AlgebraOfGraphics.data(df) * visual(BoxPlot) * 
    mapping(:generator, :value, row=:variable, col=:model, color=:generator)
plt = draw(
    plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=150, height=120), 
    facet=(; linkyaxes=:none)
)   
display(plt)
save(joinpath(output_images_path, "circles_benchmark.png"), plt, px_per_unit=5)
```

>>>>

```{julia}
include("$(pwd())/notebooks/setup.jl")
eval(setup_notebooks)
```

# Linearly Separable Data

@@

# Data:
test_size = 0.2
n_obs = Int(1000 / (1.0 - test_size))
counterfactual_data, test_data = train_test_split(
    load_blobs(n_obs; cluster_std=0.1, center_box=(-1. => 1.)); 
    test_size=test_size
)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
output_dim = length(unique(labels))
```

@@

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=batch_size,
)
α = [1.0,1.0,1e-1]      # penalty strengths
```

@@

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(model; mod_name=mod_name) for (mod_name, model) in models)
    Serialization.serialize(joinpath(output_path,"linearly_separable_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"linearly_separable_models.jls"))
end
```

@@

measure = Dict(
    :f1score => multiclass_f1score, 
    :acc => accuracy, 
    :precision => multiclass_precision
)
model_performance = DataFrame()
for (mod_name, model) in model_dict
    # Test performance:
    _perf = CounterfactualExplanations.Models.model_evaluation(model, test_data, measure=collect(values(measure)))
    _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
    _perf.mod_name .= mod_name
    model_performance = vcat(model_performance, _perf)
end
Serialization.serialize(joinpath(output_path,"linearly_separable_model_performance.jls"), model_performance)
CSV.write(joinpath(output_path, "linearly_separable_model_performance.csv"), model_performance)
model_performance
```

```{julia}
n_regen = 200
n_each = batch_size
for (mod_name, model) in model_dict
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
    opt = ImproperSGLD()
    plts = []
    for target in levels(labels)
        target_idx = findall(levels(labels) .== target)[1]
        f(x) = logits(model, x)
        X̂ = sampler(f, opt; niter=n_regen, n_samples=n_each, y=target_idx)
        ex = extrema(hcat(MLJFlux.reformat(X),X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=10)
        x2 = range(1.0f0.*ylims...,length=10)
        p(x) = probs(model, x)
        plt = Plots.contour(
            x1, x2, (x, y) -> p([x, y][:,:])[target_idx], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(
            MLJFlux.reformat(X)[1,:], MLJFlux.reformat(X)[2,:], 
            color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5
        )
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target_idx], size(X̂,2)), 
            group=repeat([target_idx], size(X̂,2)), 
            shape=:star5, ms=10
        )
        savefig(plt, joinpath(output_images_path, "linearly_separable_generated_$(mod_name).png"))
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400), plot_title=mod_name)
    display(plt)
end
```

@@

```{julia}
Random.seed!(2023)

M = model_dict["JEM"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[1]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[2]
factual = predict_label(M, counterfactual_data, x_factual)[1]

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(ce, title=name, alpha=0.2, cbar=false, axis=nothing)
    if contains(name, "ECCCo")
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.5
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(650,500))
display(plt)
savefig(plt, joinpath(output_images_path, "linearly_separable_poc.png"))
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="Linearly Separable",
            n_individuals=25,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
CSV.write(joinpath(output_path, "linearly_separable_benchmark.csv"), bmk())
```

```{julia}
df = @chain bmk() begin
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
    @mutate(variable = ifelse.(variable .== "distance", "Cost", variable))
    @mutate(variable = ifelse.(variable .== "redundancy", "Redundancy", variable))
    @mutate(variable = ifelse.(variable .== "Validity", "Validity", variable))
end
plt = AlgebraOfGraphics.data(df) * visual(BoxPlot) * 
    mapping(:generator, :value, row=:variable, col=:model, color=:generator)
plt = draw(
    plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=150, height=120), 
    facet=(; linkyaxes=:none)
)   
display(plt)
save(joinpath(output_images_path, "linearly_separable_benchmark.png"), plt, px_per_unit=5)
```

### 2023-05-09 notebooks/mnist.qmd
more results

# Target:
factual_label = 8
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 3
factual = predict_label(M, counterfactual_data, x_factual)[1]
γ = 0.9

# Training params:
T = 100
opt = Flux.Optimise.Adam(0.1)
```

```{julia}
# Search:
generic_generator = GenericGenerator(opt=Flux.Optimise.Descent(1.0))
ce_wachter = generate_counterfactual(
    x_factual, target, counterfactual_data, M, generic_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)
greedy_generator = GreedyGenerator(η=2.0)
ce_jsma = generate_counterfactual(
    x_factual, target, counterfactual_data, M, greedy_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)
```

@@

```{julia}
# Define generator:
revise_generator = REVISEGenerator(
  opt = opt,
  λ=0.0
)
# Generate recourse:
counterfactual_data.generative_model = vae # assign generative model
ce_strong = generate_counterfactual(
    x_factual, target, counterfactual_data, M, revise_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
counterfactual_data_weak = deepcopy(counterfactual_data)
counterfactual_data_weak.generative_model = vae_weak
ce_weak = generate_counterfactual(
    x_factual, target, counterfactual_data_weak, M, revise_generator;
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
```

@@

```{julia}
# JEM parameters:
𝒟x = Uniform(-1.0,1.0)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,1e-1]      # penalty strengths
```

@@

# Joint Energy Model:
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    epochs=epochs,
    batch_size=batch_size,
    finaliser=_finaliser,
    loss=_loss,
    jem_training_params=(
        α=α,verbosity=10,
    ),
    sampling_steps=20,
)

@@

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.1,0.1,0.1],
    temp=0.1,η=0.01,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = Flux.Optimise.Adam(η),
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
    T::Int = 100,
)

@@

```{julia}
plt, eccco_generator = _plot_eccco_mnist(opt=JSMADescent(η=1.0))
display(plt)
savefig(plt, joinpath(output_images_path, "mnist_eccco.png"))
```

@@

```{julia}
function plot_mnist(
    factual::Int,target::Int;
    generator::AbstractGenerator,
    model::AbstractFittedModel=model_dict["JEM Ensemble"],
    data::CounterfactualData=counterfactual_data,
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
    _plot_title::Bool=true,
    kwargs...,
)
    decision_threshold = !isdefined(kwargs, :decision_threshold) ? 0.9 : decision_threshold
    max_iter = !isdefined(kwargs, :max_iter) ? 100 : max_iter
    initialization = !isdefined(kwargs, :initialization) ? :identity : initialization
    converge_when = !isdefined(kwargs, :converge_when) ? :generator_conditions : converge_when

@@

    _x = CounterfactualExplanations.counterfactual(ce)
    plt = Plots.plot(
        convert2image(MNIST, reshape(_x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    return plt
end
```

```{julia}
if _regen
    function plot_all_digits(rng=1;verbose=true,kwargs...)
        plts = []
        for i in 0:9
            for j in 0:9
                @info "Generating counterfactual for $(i) -> $(j)"
                plt = plot_mnist(i,j;kwargs...,rng=rng)
                !verbose || display(plt)
                plts = [plts..., plt]
            end
        end
        plt = Plots.plot(plts...; size=(img_height*10,img_height*10), layout=(10,10))
        return plt
    end
    plt = plot_all_digits(generator=eccco_generator)
    savefig(plt, joinpath(output_images_path, "mnist_eccco_all_digits.png"))
end
```

### Different Generators

```{julia}
# Setup:
model = model_dict["JEM Ensemble"]

# Benchmark generators:
generator_dict = Dict(
    :wachter => generic_generator,
    :revise => revise_generator,
    :greedy => greedy_generator,
    :eccco => eccco_generator,
)

ces = Dict()
for (gen_name, gen) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, model, gen; 
        decision_threshold=γ, max_iter=T,
        initialization=:identity,
        converge_when=:generator_conditions,
    )
    ces[gen_name] = ce
end
plt_order = sortperm(collect(keys(ces)))

# Plot:
p1 = Plots.plot(
    convert2image(MNIST, reshape(x_factual,28,28)),
    axis=nothing, 
    size=(img_height, img_height),
    title="Factual"
)

plts = []
for (_name,ce) in ces
    _x = CounterfactualExplanations.counterfactual(ce)
    _phat = target_probs(ce)
    _title = "$_name (p̂=$(round(_phat[1]; digits=3)))"
    plt = Plots.plot(
        convert2image(MNIST, reshape(_x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    plts = [plts..., plt]
end
plts = plts[plt_order]
plts = [p1, plts...]
plt = Plots.plot(plts...; size=(img_height*length(plts),img_height), layout=(1,length(plts)))
display(plt)
savefig(plt, joinpath(output_images_path, "mnist_all_generators.png"))
```

@@

@@

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="MNIST",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
)
CSV.write(joinpath(output_path, "mnist_benchmark.csv"), bmk())
```

```{julia}
@chain bmk() begin
    @group_by(dataname, generator, model, variable)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable == "distance_from_energy")
end
```

>>>>

# Target:
factual_label = 9
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 7
factual = predict_label(M, counterfactual_data, x_factual)[1]
γ = 0.9

# Training params:
T = 100
```

```{julia}
# Search:
generic_generator = WachterGenerator()
ce_wachter = generate_counterfactual(
    x_factual, target, counterfactual_data, M, generic_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)
greedy_generator = GreedyGenerator(η=2.0)
ce_jsma = generate_counterfactual(
    x_factual, target, counterfactual_data, M, greedy_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)
```

@@

```{julia}
# Define generator:
revise_generator = REVISEGenerator(
    opt = Flux.Optimise.Descent(0.25),
    λ=0.0,
)
# Generate recourse:
counterfactual_data.generative_model = vae # assign generative model
ce_strong = generate_counterfactual(
    x_factual, target, counterfactual_data, M, revise_generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
counterfactual_data_weak = deepcopy(counterfactual_data)
counterfactual_data_weak.generative_model = vae_weak
ce_weak = generate_counterfactual(
    x_factual, target, counterfactual_data_weak, M, revise_generator;
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
```

@@

```{julia}
# JEM parameters:
𝒟x = Uniform(-1.0,1.0)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,2e-2]      # penalty strengths
```

@@

# Joint Energy Model:
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    epochs=epochs,
    batch_size=batch_size,
    finaliser=_finaliser,
    loss=_loss,
    jem_training_params=(
        α=α,verbosity=10,
    ),
    sampling_steps=25,
)

@@

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.1,0.25,0.25],
    temp=0.1,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = nothing,
    rng::Union{Int,AbstractRNG}=1234,
    T::Int = 100,
)

@@

```{julia}
plt, eccco_generator = _plot_eccco_mnist()
display(plt)
savefig(plt, joinpath(output_images_path, "mnist_eccco.png"))
```

@@

```{julia}
function plot_mnist(
    factual::Int,target::Int;
    generator::AbstractGenerator,
    model::AbstractFittedModel=model_dict["JEM Ensemble"],
    data::CounterfactualData=counterfactual_data,
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
    _plot_title::Bool=true,
    show_factual::Bool=false,
    kwargs...,
)
    Random.seed!(rng)

    decision_threshold = !isdefined(kwargs, :decision_threshold) ? 0.9 : decision_threshold
    max_iter = !isdefined(kwargs, :max_iter) ? 100 : max_iter
    initialization = !isdefined(kwargs, :initialization) ? :identity : initialization
    converge_when = !isdefined(kwargs, :converge_when) ? :generator_conditions : converge_when

@@

    _x = CounterfactualExplanations.counterfactual(ce)
    plt = Plots.plot(
        convert2image(MNIST, reshape(_x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    if show_factual
        plt_factual = Plots.plot(
            convert2image(MNIST, reshape(x,28,28)),
            axis=nothing, 
            size=(img_height, img_height),
            title="Factual"
        )
        plt = Plots.plot(plt_factual, plt; size=(img_height*2,img_height), layout=(1,2))
    end

    return plt
end
```

```{julia}
if _regen
    function plot_all_digits(rng=123;verbose=true,kwargs...)
        plts = []
        for i in 0:9
            for j in 0:9
                @info "Generating counterfactual for $(i) -> $(j)"
                plt = plot_mnist(i,j;kwargs...,rng=rng)
                !verbose || display(plt)
                plts = [plts..., plt]
            end
        end
        plt = Plots.plot(plts...; size=(img_height*10,img_height*10), layout=(10,10))
        return plt
    end
    plt = plot_all_digits(generator=eccco_generator)
    savefig(plt, joinpath(output_images_path, "mnist_eccco_all_digits.png"))
end
```

@@

```{julia}
λ₁ = 0.1
λ₂ = 0.25
λ₃ = 0.25
Λ = [λ₁, λ₂, λ₃]

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(λ=λ₁),
    "REVISE" => REVISEGenerator(λ=λ₁),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
)
```

@@

bmks = []
for target in sort(unique(labels))
    for factual in sort(unique(labels))
        if factual == target
            continue
        end
        bmk = benchmark(
            counterfactual_data; 
            models=model_dict, 
            generators=generator_dict, 
            measure=measures,
            suppress_training=true, dataname="MNIST",
            n_individuals=5,
            target=target, factual=factual,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)

CSV.write(joinpath(output_path, "mnist_benchmark.csv"), bmk())
```
